Randomization Examples using `nycflights13` `flights` data

### 2017-12-27 inst/doc/flights_examples.Rmd
initial commit

---
title: "Randomization Examples using `okcupiddata` `profiles` data"
author: "Chester Ismay"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    df_print: kable
vignette: |
  %\VignetteIndexEntry{OKCupid profiles example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r include=FALSE}
knitr::opts_chunk$set(fig.width = 8, fig.height = 5, cache = TRUE) 
```

@@

```{r message=FALSE, warning=FALSE}
library(okcupiddata)
library(stringr)
library(infer)
set.seed(2017)
prof_small <- profiles %>% 
  dplyr::sample_n(size = 500) %>% 
  dplyr::mutate(city = dplyr::case_when(
    str_detect(location, "san fran") ~ "san fran",
    !str_detect(location, "san fran") ~ "not san fran"
  )) %>% 
  dplyr::select(age, sex, city, drugs, height, status)
```

- `height` and `age` are numerical variables.
- `sex` has two categories (`"m"`, `"f"`)
- `city` has two categories (`"san fran"`, `"not san fran"`)
- `drugs` has three categories (`"never"`, `"sometimes"`, `"often"`) - Also has missing values
- `status` has three categories (`"single"`, `"available"`, `"seeing someone"`)

@@

One numerical variable (mean)

```{r}
prof_small %>%
  specify(response = age) %>% # alt: age ~ NULL (or age ~ 1)
  hypothesize(null = "point", mu = 50) %>% 
  generate(reps = 1000, type = "bootstrap") %>% 
  calculate(stat = "mean") %>% 
  visualize()
```

One numerical variable (median)

```{r}
prof_small %>%
  specify(response = age) %>% # alt: age ~ NULL (or age ~ 1)
  hypothesize(null = "point", med = 55) %>% 
  generate(reps = 1000, type = "bootstrap") %>% 
  calculate(stat = "median") %>% 
  visualize()
```

One categorical (2 level) variable

```{r}
prof_small %>%
  specify(response = sex) %>% # alt: sex ~ NULL (or sex ~ 1)
  hypothesize(null = "point", p = c("m" = .65)) %>% 
  generate(reps = 1000, type = "simulate") %>% 
  calculate(stat = "prop") %>% 
  visualize()
```

@@

@@

One numerical variable one categorical (2 levels) (diff in medians)

>>>>

---
title: "Randomization Examples using `nycflights13` `flights` data"
author: "Chester Ismay and Andrew bray"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    df_print: kable
vignette: |
  %\VignetteIndexEntry{flights example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r include=FALSE}
knitr::opts_chunk$set(fig.width = 8, fig.height = 5) 
```

@@

```{r message=FALSE, warning=FALSE}
library(nycflights13)
library(dplyr)
library(ggplot2)
library(stringr)
library(infer)
set.seed(2017)
fli_small <- flights %>% 
  sample_n(size = 500) %>% 
  mutate(season = case_when(
    month %in% c(10:12, 1:3) ~ "winter",
    month %in% c(4:9) ~ "summer"
  )) %>% 
  mutate(day_hour = case_when(
    between(hour, 1, 12) ~ "morning",
    between(hour, 13, 24) ~ "not morning"
  )) %>% 
  select(arr_delay, dep_delay, season, 
         day_hour, origin, carrier) %>%
  filter(., complete.cases(.))
```

* Two numeric - `arr_delay`, `dep_delay`
* Two categories 
    - `season` (`"winter"`, `"summer"`), 
    - `day_hour` (`"morning"`, `"not morning"`)
* Three categories - `origin` (`"EWR"`, `"JFK"`, `"LGA"`)
* Sixteen categories - `carrier`

@@

### One numerical variable (mean)

```{r}
x_bar <- fli_small %>%
  summarize(mean(dep_delay)) %>%
  pull()
null <- fli_small %>%
  specify(response = dep_delay) %>%
  hypothesize(null = "point", mu = 10) %>% 
  generate(reps = 1000, type = "bootstrap") %>% 
  calculate(stat = "mean")
ggplot(null, aes(x = stat)) +
  geom_density() +
  geom_vline(xintercept = x_bar, col = "red")
null %>%
  summarize(mean(stat > x_bar)) * 2
```

### One numerical variable (median): todo

### One categorical (2 level) variable

```{r}
p_hat <- fli_small %>%
  summarize(mean(day_hour == "morning")) %>%
  pull()
null <- fli_small %>%
  specify(response = dep_delay, success = "morning") %>%
  hypothesize(null = "point", p = .5) %>% 
  generate(reps = 1000, type = "simulate") %>% 
  calculate(stat = "prop")
ggplot(null, aes(x = stat)) +
  geom_density() +
  geom_vline(xintercept = p_hat, col = "red")
null %>%
  summarize(mean(stat > p_hat)) * 2
```

@@

One numerical variable, one categorical (2 levels) (diff in means)

@@

One numerical variable, one categorical (2 levels) (diff in medians)
